---
title: "p8105_hw6_sd3730"
author: "Stacey Dai"
date: "2024-12-02"
output: github_document
---

# Problem 1

Load necessary packages for problem 1.
```{r p1setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(rnoaa)
library(broom)
```

Clean and prepare the noaa dataset for analysis.
```{r p1cleanup, message = FALSE, warning = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31"
  ) |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10
  ) |>
  select(name, id, everything())
```

Generate 5000 bootstrap samples and fit a simple linear regression for each sample.
```{r bootstrap, message = FALSE, warning = FALSE}
set.seed(123)

bootstrap_results = 
  replicate(
    5000, 
    {
      sample_data = weather_df |> slice_sample(prop = 1, replace = TRUE)
      model = lm(tmax ~ tmin, data = sample_data)
      
      r2 = glance(model)$r.squared
      coefs = tidy(model)
      log_product = log(coefs$estimate[1] * coefs$estimate[2])
      
      tibble(r2 = r2, log_product = log_product)
    },
    simplify = FALSE
  ) |>
  bind_rows()
```

Identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval.
```{r 95ci, message = FALSE, warning = FALSE}
ci_r2 = quantile(bootstrap_results$r2, probs = c(0.025, 0.975))
ci_log_product = quantile(bootstrap_results$log_product, probs = c(0.025, 0.975))
```

Plot histograms for both quantiles.
```{r p1plot, message = FALSE, warning = FALSE}
bootstrap_results |>
  pivot_longer(cols = everything(), names_to = "metric", values_to = "value") |>
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black", alpha = 0.7) +
  facet_wrap(~metric, scales = "free", ncol = 1) +
  theme_minimal() +
  labs(
    title = "Bootstrap Distributions",
    x = "Estimated Value",
    y = "Frequency"
  )
```

The distribution for the log_product graph appears to be symmetric andd approximately normal. The distribution for the r2 graphh is symmetrical and unimodal. 

# Problem 2

Load necessary packages for problem 2.
```{r p2setup, message = FALSE, warning = FALSE}
library(purrr)
```

Prepare and clean data for analysis.
```{r p2cleanup, message = FALSE, warning = FALSE}
homicides = read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")

homicides_clean = homicides |>
  mutate(
    city_state = str_c(city, ", ", state),
    solved_binary = if_else(disposition == "Closed by arrest", 1, 0),
    victim_age = suppressWarnings(as.numeric(victim_age))
  ) |>
  filter(
    !is.na(victim_age),
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")),
    victim_race %in% c("White", "Black")
  )
```

Model the binary outcome of solving homicides in Baltimore, MD.
```{r baltimore_analysis, echo = TRUE, message = FALSE, warning = FALSE}
baltimore_df = homicides_clean |>
  filter(city_state == "Baltimore, MD")

baltimore_model = glm(
  solved_binary ~ victim_age + victim_sex + victim_race,
  data = baltimore_df,
  family = binomial()
)

baltimore_results = broom::tidy(baltimore_model, exponentiate = TRUE, conf.int = TRUE)

male_vs_female_or = baltimore_results |>
  filter(term == "victim_sexMale") |>
  select(estimate, conf.low, conf.high)

print(baltimore_results)
print(male_vs_female_or)
```

Next, fit a logistic regression for all cities.
```{r allcities, message = FALSE, warning = FALSE}
city_results = homicides_clean |>
  nest(data = -city_state) |>
  mutate(
    model = map(data, ~ glm(solved_binary ~ victim_age + victim_sex + victim_race, 
                            data = ., family = binomial())),
    results = map(model, ~ broom::tidy(., exponentiate = TRUE, conf.int = TRUE))
  ) |>
  unnest(results) |>
  filter(term == "victim_sexMale") |>
  select(city_state, estimate, conf.low, conf.high)

print(city_results)
```

Create a plot with the estimated ORs and CIs for each city.
```{r p2plot, message = FALSE, warning = FALSE}
city_results_sorted = city_results |>
  arrange(estimate) |>
  mutate(city_state = fct_reorder(city_state, estimate))

ggplot(city_results_sorted, aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides (Male vs Female Victims)",
    x = "City",
    y = "Adjusted Odds Ratio"
  )
```


